---
title: "Skills Quiz: Lowess Curves and Weighted Regressions"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    toc: true
    toc_float: true
---


## Instructions

Use this file to keep a record of your work as you complete the "Skills Quiz: Lowess Curves and Weighted Regressions" assignment in Canvas.


```{r, message=FALSE, warning=FALSE}
library(mosaicData)
library(ggplot2)
```


----

<!-- Note: The {} after each Problem and Part header allows you to keep track of what work you have completed. Write something like {Done} once you complete each problem and your html file will then show you a nice summary of what you have "done" already. -->



## Problem 1 {}

Run this code in R. Notice that the loess curves being fit to the data do not "fit" very well currently. Work through the following parts of this problem to learn how to improve the fit of a loess curve.

```{r, echo=TRUE}
#Ensure library(ggplot2) and library(mosaicData) are loaded
ggplot(Births78, aes(x=day_of_year, y=births, color=wday)) +
  geom_point() +
  geom_smooth(method="loess", span=0.75,
              se=FALSE, 
              method.args=list(degree=1))
```

### Part (a)

Copy and paste the code from above, then change the `span` and the `degree` until you come up with a nice fitting set of loess curves. Write a brief comment about how these two choices effect the fit of your loess curves.

* `span` adjusts the percentage of points that are used in each local regression. It should be a number between 0 and 1. 

* `degree` changes the type of polynomial that is used in each local regression. A choice of `degree = 1` uses lines, `degree = 2` uses parabolas, `degree = 3` uses cubic functions, and so on.

Hint: it may help to add `+ facet_wrap(~wday)` to the end of the above code, which will allow you to see each loess curve in its own little plot.

<div class="YourAnswer">

```{r}
#Type your code here...
```

Write your comments here...

</div>

### Part (b)

Use your nicely fitting loess curve that you came up with in Part (a) to predict the number of births that happened on day 366 of 1978... (i.e., January 1st of 1979). How accurate do you think this prediction will be? Why?

Hint: `View(Births78)` and recall that 366 comes after 365...

<div class="YourAnswer">

```{r}
#Hint:
#mylo <- loess(Y ~ X1 + X2, data=..., span=..., degree=...)
#predict(mylo, data.frame(X1=..., X2=...))
```

</div>


## Problem 2 {}

Open the `CO2` data set in R. This data is based around an experiment aimed to understand CO2 `uptake` of plants, which is an important characteristic of a plant's overall health and longevity. Variables known to be associated with `uptake` were controlled in this experiment at different values to try to model their effect on `uptake`. 

* `conc` - the ambient CO2 concentration surrounding the plant. The more CO2 available, the higher the uptake levels should be.

* `Type` - The type of plant. Plants from the warmer climate of Mississippi were compared to plants from a colder climate like Quebec.

* `Treatment` - Plants were either chilled or left un-chilled to see how refrigeration preserved CO2 uptake.


### Part (a)

Create a scatterplot with overlaid loess curves showing how uptake is effected by the (1) ambient CO2 `conc`entration, (2) the `Type` of plant, and (3) the `Treatment` applied to the plant. 

Fit loess curves to the data using span=0.6 and degree=2. (You should have four different curves.)

Hint: use a ggplot and `color=interaction(Type, Treatment)`

<div class="YourAnswer">

```{r}
#Type your code here...
```

Write your answer here...

</div>

### Part (b)

Using one of the four loess curves that were drawn on your plot, predict the `uptake` of CO2 for a plant that is chilled, from Quebec, and is placed in an ambient CO2 concentration of 300 mL/L.

Note: It would be awesome if you could run something like `loess(uptake ~ conc + Treatment + Type, data=CO2)`. Unfortunately, this will not work. Instead, you will need to run something like `loess(uptake ~ conc, data= ...)` where the `...` is an appriopriately reduced version of the `CO2` data. Also, be sure to specify the span and degree in your loess(...) function as instructed in Part (a).


<div class="YourAnswer">

```{r}
#Type your code here...
```

</div>

<br/>


## Problem 3 {}

Open the `Loblolly` data set in R. Sometimes a lowess curve can be used to help us decide on an appropriate model for data.

### Part (a)

Create a scatterplot of the `height` of a tree on the `age` of the tree. Overlay both a simple linear regression line and a lowess curve, degree 1 on the plot. Does the lowess curve suggest that the line is a good fit or a poor fit?

<div class="YourAnswer">

```{r}
#Type your code here...
```

Type your comments here...

</div>


### Part (b)

Re-draw the scatterplot, this time overlaying the curve from a quadratic regression model as well as the lowess curve of degree 1. Does this model look better or worse than the simple linear regression line when compared to the lowess curve?

<div class="YourAnswer">

```{r}
#Type your code here...
```

Type your comments here...

</div>



## Problem 4 {}

When regressions are in low dimensional space, like 2D or 3D, we can often visually identify problems with outliers. But identifying outliers becomes more difficult in higher dimensions. Weighted regressions, which unlike loess curves work in any dimensional space, are a powerful way to overcome this issue.

Open the `Davis` data set from `library(car)` in R.

### Part (a)

Create a scatterplot of weight on height using the `Davis` data set. Overlay a simple linear regression line and a lowess (or loess) curve of your choice. 

Notice the extreme outlier. How is it effecting both the regression line and the lowess (or loess) curve?

<div class="YourAnswer">

```{r}
#Type your code here...
```

Type your comments here...

</div>


### Part (b)

Perform a weighted regression of weight on height where the weights are defined as 1/abs(residuals from the simple linear regression of Part a). Plot both the original regression, the lowess curve, and this weighted regression on the same plot. 

Does this improve the fit of the regression line? 

What is the equation of this weighted regression line?

<div class="YourAnswer">

```{r}
#Type your code here...
```

Type your comments here...

</div>

### Part (c)

Perform four more weighted regressions (making five total weighted regressions) of weight on height. In each case, let the weights be defined as 1/abs(residuals from the previous weighted regression). Plot the original simple linear regression, the lowess curve, and each of your weighted regressions on the same plot. Where does the weighted regression line seem to be converging? 

<div class="YourAnswer">

```{r}
#Type your code here...
```

Type your comments here...

</div>


### Part (d)

Load `library(MASS)`. Then use rlm(...) instead of lm(...) to perform what is called a "robust" simple linear regression of weight on height. Do not use weighting this time. This rlm(...) function does essentially what you were doing in Part (c), it continues to perform weighted regressions until the line "converges" to a stable slope and intercept, but the rlm(...) function only provides you with the final results. (Which is nice.) 

Create a scatterplot that shows: (1) your original simple linear regression, (2) your lowess (or loess) curve, (3) all of your five weighted regression lines, and (4) the robust regression line. Comment on what you notice about the relationship of all of these "curves". 

<div class="YourAnswer">

```{r}
#Type your code here...
```

Type your comments here...

</div>


### Part (e)

Use your robust regression model to predict the weight of an individual who is 140 cm tall. Similarly, use your original simple linear regression to make the same prediction. How do these two predictions compare?

<div class="YourAnswer">

```{r}
#Type your code here...
```

Type your comments here...

</div>











<style>

.YourAnswer {
  color: #317eac;
  padding: 10px;
  border-style: solid;
  border-width: 2px;
  border-color: skyblue4;
  border-radius: 5px;
}

</style>

 
 