---
title: "Skills Quiz: Different Types of Models"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    toc: true
    toc_float: true
---


## Instructions

Use this file to keep a record of your work as you complete the "Skills Quiz: Different Types of Models" assignment in Canvas.


```{r, message=FALSE, warning=FALSE}
library(mosaicData)
```


----

<!-- Note: The {} after each Problem and Part header allows you to keep track of what work you have completed. Write something like {Done} once you complete each problem and your html file will then show you a nice summary of what you have "done" already. -->



## Problem 1 {}

Consider the scatterplot shown here of a single residence's monthly gas bill according to the month of the year. See `?Utilities` for more details on the data.

```{r}
plot(gasbill ~ month, data=Utilities, main="Single Residence in Minnisota", xlab="Month of the Year", ylab="Monthly Gas Bill (US Dollars)")


```



### Part (a)

Add the estimated regression function to the scatterplot above and state the function below.

<div class="YourAnswer">

$$
  \hat{Y}_i = ...
$$

</div>

### Part (b)

Diagnose the appropriateness of this regression model. How well does it fit the data? 

<span style="font-size:.8em;">Be sure to provide diagnostic plots and supporting arguments for your claims.</span>

<div class="YourAnswer">

```{r}
#Type your code here...
```

</div>


### Part (c)

What range of possible gas bill amounts do you predict for the September bill? How confident are you in your prediction?

<div class="YourAnswer">

```{r}
#Type your code here...
```

Type your answer here...

</div>


### Part (d)

Compute the mean of just September's gas bills and draw a horizontal reference line for this mean on your scatterplot. How does this mean compare to your prediction in Part (c)?

<div class="YourAnswer">

```{r}
#Type your code here...
```

Type your answer here...

</div>




### Part (e)

State the values of $R^2$, $MSE$, and the residual standard error. What do each of these values tell us about the regression model?

<div class="YourAnswer">

```{r}
#Type your code here...
```

Type your answer here...

</div>

<br />



## Problem 2

View the `mtcars` dataset and corresponding help file `?mtcars`.

Perform a regression that predicts the miles per gallon `mpg` of the vehicle based on the quarter mile time `qsec` and transmission type `am` of the vehicle.


### Part (a)

Plot the data and your fitted regression model.

<div class="YourAnswer">

```{r}
#Type your code here...
```

</div>


### Part (b)

State the fitted regression model.

<div class="YourAnswer">

$$
 ...
$$

</div>

### Part (c)

Use `summary(...)` to perform an appropriate t test to determine if the interaction term is needed in this regression model.

<div class="YourAnswer">

```{r}
#Type your code here...
```

Type your answer here...

</div>


### Part (d)

Diagnose the appropriateness of this regression model. How well does it fit the data? 

<span style="font-size:.8em;">Be sure to provide diagnostic plots and supporting arguments for your claims.</span>

<div class="YourAnswer">

```{r}
#Type your code here...
```

Type your answer here...

</div>


### Part (e)

State the $R^2$ and residual standard error values for this model. What do these values show?

<div class="YourAnswer">

```{r}
#Type your code here...
```

Type your answer here...

</div>

<br />



## Problem 3

View the `mtcars` dataset and corresponding help file `?mtcars`.

Perform a regression that predicts the quarter mile time `qsec` of the vehicle based on the displacement of the engine `disp` and transmission type `am` of the vehicle according to the model:

$$
  \underbrace{Y_i}_\text{qsec} = \beta_0 + \beta_1 \underbrace{X_{1i}}_\text{disp} + \beta_2 X_{1i}^2 + \beta_3 \underbrace{X_{2i}}_\text{am == 1} + \beta_4 X_{1i}X_{2i} + \beta_5 X_{1i}^2X_{2i} + \epsilon_i
$$

where $\epsilon_i \sim N(0, \sigma^2)$.

### Part (a)

Plot the data and your fitted regression model.

<div class="YourAnswer">

```{r}
#Type your code here...
```

Type your answer here...

</div>

### Part (b)

State the fitted regression model.

<div class="YourAnswer">

$$
  \hat{Y}_i = ...
$$

</div>


### Part (c)

Use `summary(...)` to perform appropriate t tests to determine which interaction terms are needed in this regression model.

<div class="YourAnswer">

```{r}
#Type your code here...
```

</div>

### Part (d)

Diagnose the appropriateness of this regression model. How well does it fit the data? 

<span style="font-size:.8em;">Be sure to provide diagnostic plots and supporting arguments for your claims.</span>


<div class="YourAnswer">

```{r}
#Type your code here...
```

Type your answer here...

</div>

### Part (e) 

Drop the "least significant" term from the model. This is a silly statement really, but is an accepted practice when searching for a "best" model in regression analysis. What has changed?

<div class="YourAnswer">

```{r}
#Type your code here...
```

Type your answer here...

</div>










<style>

.YourAnswer {
  color: #317eac;
  padding: 10px;
  border-style: solid;
  border-width: 2px;
  border-color: skyblue4;
  border-radius: 5px;
}

</style>

 
 