---
title: "Skills Quiz: Residuals, Sums of Squares, and R-squared"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    toc: true
    toc_float: true
---


## Instructions

Use this file to keep a record of your work as you complete the "Skills Quiz: Residuals, Sums of Squares, and R-squared" assignment in Canvas.


----

<!-- Note: The {} after each Problem and Part header allows you to keep track of what work you have completed. Write something like {Done} once you complete each problem and your html file will then show you a nice summary of what you have "done" already. -->

## Problem 1 {}

Open the `Orange` dataset in R. As stated in the help file for this data set, "The Orange data... records of the growth of orange trees." 

Perform a simple linear regression of the circumference of the tree based on its age in days.

### Part (a) {}

Type out the mathematical equation for this regression model and label both $Y$ and $X$ in the equation.

<div class="YourAnswer">

$$
  \underbrace{Y_i}_\text{some label} = ...
$$

</div>


### Part (b) {}
 
Plot a scatterplot of the data with your regression line overlaid.

<div class="YourAnswer">

```{r}
# Type your code here
```

</div>


### Part (c) {}

State and interpret the slope and y-intercept of this regression. Are they meaningful for this data?

<div class="YourAnswer">

Type your answer here...

</div> 
 
 
### Part (d) {}

What are the values of SSE, SSR, SSTO, $R^2$, and the correlation $r$ for this regression?
 
<div class="YourAnswer">

```{r}
# Type your code here
```

</div>


### Part (e) {}

How do the values of SSR and SSTO relate to the correlation?

<div class="YourAnswer">

Type your answer here...

</div>


### Part (f) {}

What circumference would we expect orange trees to have on average after 3 years, based on this regression? 

<div class="YourAnswer">

Type your answer here...

</div>



----

## Problem 2 {}

Open the **mtcars** dataset in R. Fit three different regression models to the data that can each be used to explain average gas mileage of a vehicle (`mpg`). 

* The first regression should use the weight (`wt`) of the vehicle as the explanatory variable.
* The second should use the number of cylinders (`cyl`) of the engine of the vehicle as the explanatory variable.
* The third should use the gross horsepower of the vehicle (`hp`) as the explanatory variable.

### Part (a) {}

Plot all three regressions in three separate plots.

<div class="YourAnswer">

```{r}
# Type your code here..
```

</div>


### Part (b) {}

State the values of SSE, SSR, SSTO, and $R^2$ for each regression. 

<div class="YourAnswer">

```{r}
# Type your code here
```


</div>


### Part (c) {}

Compare the values from **Part (b)** across each regression. Consider these questions as you compare these numbers.

* What insight do these numbers give about the regression?
* Which numbers change, which stay the same, in these regressions? Why?
* Which regression is best at explaining average `mpg` according to these numbers?

<div class="YourAnswer">

Type your answer here...

</div> 
 
 



----

Before we can really trust the $R^2$ value from a regression model, there are important diagnostic checks to perform on the regression. 

You now have a better understanding of what a **residual** and a **fitted-value** are in regression. With that improved knowledge, work through the following problem.



----

## Problem 3 {}

Check the following technical details for each of the three regressions of **Problem 2**.

### Part (a) {}

Create a (1) residuals vs. fitted-values plot, (2) Q-Q Plot of the residuals, and (3) residuals vs. order plot for each regression of **Problem 2**.

<div class="YourAnswer">

```{r}
# Type your code here
```

</div>


### Part (b) {}

Explain, as best you understand currently, what each of these three plots show for these regressions.

<div class="YourAnswer">

Type your answer here...

</div>


### Part (c) {}

Comment on whether or not we should trust the $R^2$ value from each of your three regressions based on your plots in Part (a). 

<div class="YourAnswer">

Type your answer here...

</div>



----











<style>

.YourAnswer {
  color: #317eac;
  padding: 10px;
  border-style: solid;
  border-width: 2px;
  border-color: skyblue4;
  border-radius: 5px;
}

</style>

 
 